Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-09-04 papers/integration/body_text.tex
The body_text.tex file that I uploaded to the Frontiers web interface

% While writing, don't stop for errors
\nonstopmode

% The Frontiers template.
\documentclass{frontiersSCNS}

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree

\linenumbers

% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
% End language alias trick

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{James {et~al.}}
\def\Authors{Sebastian James\,$^{1,2,*}$, Chris Papapavlou\,$^5$, Alexander Blenkinsop\,$^{1,2}$, \\
  Alexander Cope\,$^3$,  \\
  Sean Anderson\,$^{2,4}$, Konstantinos Moustakas\,$^5$, Kevin Gurney\,$^{1,2}$}
% Affiliations should be keyed to the author's name with superscript
% numbers and be listed as follows: Laboratory, Institute, Department,
% Organization, City, State abbreviation (USA, Canada, Australia), and
% Country (without detailed address information such as city zip codes
% or street names).
%
% If one of the authors has a change of address, list the new address
% below the correspondence details using a superscript symbol and use
% the same symbol to indicate the author in the author list.
\def\Address{\\
$^{1}$Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
$^{2}$Insigneo Institute for in-silico Medicine,
  The University of Sheffield, Sheffield, UK \\
$^{3}$Department of Computer Science,
  The University of Sheffield, Sheffield, UK \\
$^{4}$Department of Automatic Control Systems Engineering,
  The University of Sheffield, Sheffield, UK \\
$^{5}$Department of Electrical and Computer Engineering,
  The University of Patras, Patras, Greece \\
}
% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Kevin Gurney}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{k.gurney@sheffield.ac.uk}

\begin{document}

% Aliases
\newcommand{\ccg}{Cope-Chambers-Prescott-Gurney}
\newcommand{\stob}{SpineML\_2\_BRAHMS}
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\blue}{\textcolor{blue}}
% I like my citations in blue
\newcommand{\mycite}[1]{\blue{\cite{#1}}}
% expon function is roman e
\newcommand{\expon}{\mathrm{e}}
% A font to indicate a branch of a github repository
\newcommand{\branch}[1]{\textbf{\texttt{#1}}}
\newcommand{\cmnt}[1]{\blue{#1}}
\newcommand{\dg}{$\degree$}

\onecolumn
\firstpage{1}

\title[Integrated brain and biomechanics]{
Integrating brain and biomechanical models - a new paradigm
for understanding  neuro-muscular control
}

\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{} % If more than 1 corr. author refer to original template
             % and fill this in
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% The sections below are for reference only.
%%%
%%% For Original Research Articles, Clinical Trial Articles, and
%%% Technology Reports the section headings should be those
%%% appropriate for your field and the research itself. It is
%%% recommended to organize your manuscript in the following sections
%%% or their equivalents for your field:
%%%
%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
%%%
%%% Please note that the Material and Methods section can be placed in
%%% any of the following ways: before Results, before Discussion or
%%% after Discussion.
%%%
%%% For information about Clinical Trial Registration, please go to
%%% http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
%%%
%%% For Clinical Case Studies the following sections are mandatory:
%%% Abstract, Introduction, Background, Discussion, and Concluding
%%% Remarks.
%%%
%%% For all other article types there are no mandatory sections.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\begin{abstract}

\section{}
To date, realistic models of how the central nervous system governs
behaviour have been restricted in scope to the brain, brainstem or
spinal column, as if these existed as disembodied organs.
Further, the model is often exercised in relation to an in vivo
physiological experiment with input comprising an impulse, a periodic
signal or constant activation, and output as a pattern of neural
activity in one or more neural populations. Any link to behaviour is
inferred only indirectly via these activity patterns.
%
We argue that to discover the principles of operation of neural
systems, it is necessary to express their behaviour in terms of
physical movements of a realistic motor system, and to supply inputs
that mimic sensory experience. To do this with confidence, we must
connect our brain models to neuro-muscular models and provide relevant
visual and proprioceptive feedback signals, thereby closing the loop
of the simulation.
%
This paper describes an effort to develop just such an integrated
brain and biomechanical system using a number of pre-existing
models. It describes a model of the saccadic oculomotor system
incorporating a neuromuscular model of the eye and its six extraocular
muscles. The position of the eye determines how illumination of a
retinotopic input population projects information about the location
of a saccade target into the system.
%
A pre-existing saccadic burst generator model was incorporated into
the system, which generated motoneuron activity patterns suitable for
driving the biomechanical eye. The model was demonstrated to make
accurate saccades to a target luminance under a set of environmental
constraints.
%
Challenges encountered in the development of this model showed the
importance of this integrated modelling approach. Thus, we exposed
shortcomings in individual model components which were only apparent
when these were supplied with the more plausible inputs available in a
closed loop design. Consequently we were able to suggest missing
functionality which the system would require to reproduce more
realistic behaviour.
%
The construction of such closed-loop animal models constitutes a new
paradigm of \e{computational neurobehaviour} and promises a more
thoroughgoing approach to our understanding of the brain's function as
a controller for movement and behaviour.

\tiny
 \keyFont{
   \section{Keywords:}
   integrated brain biomechanics neuromuscular oculomotor saccade basal ganglia
 } % 5 to 8 keywords
\end{abstract}

@@

% New introduction, tailored to the paradigm shift title.

% Background
The field of computational neuroscience has provided
many \e{systems models} of the brain \citep{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}.
We refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain. Whilst they differ in scale and complexity, these models
all seek to
describe the fundamental mechanisms behind common animal behaviours
such as locomotion, threat evasion, reaching or feeding. However,
none of the models cited here actually reproduce these
behaviours. In each case, the activity in a certain population of
neurons is taken to be representative of a behavioural outcome.
% Sometimes, ok to just have a brain model
In some cases, it \e{is} reasonable to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a basal
ganglia model
\citep{nambu_discharge_1990,kuhn_event-related_2004}. The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. To validate the model, the
error rates which it generates
could be compared with experimentally determined error rates in
primate subjects. We refer
to this as an \e{output assumption model} because the output is assumed
to signify behaviour. (An \e{input assumption model} assumes that sensory
input produces some particular form of neural activity in an input
population of the model.)

% Gap
However,
we may be interested in reproducing accurate simulated \e{trajectories},
in order to find out how degradation of parts of the model affect
movement. In Parkinson's Disease, degradation of the
dopamine neurons originating in the substantia nigra pars compacta (SNc)
causes diskinesia \citep{galvan_pathophysiology_2008}, as well as
abnormal network activity in
the basal ganglia \citep{brown_dopamine_2001,mccarthy_striatal_2011}.
Sufferers of the disease would be
expected to produce abnormal decision-making \e{and} movement trajectories
in a reach-to-the-correct-target task such as the one described in
\cite{james_target-distractor_2017}.
%
% Solution
A model which
sought to explore in detail the effects of the SNc degradation both on
the decision making \e{and} on the movement dynamics would need a
physically accurate virtual arm, as well as physically realistic sensory
input for the brain. This is no less than a complete model of those sections
of the brain and body which act to fulfil the task. Such a modelling
effort, if successful, would result in a virtual robot capable of expressing
behaviour \e{in response to sensory input from its environment}. This would
represent a paradigm shift in the field of computational neuroscience
worthy of the new name of \e{computational neurobehaviour}.

% Existing models don't control movement unless they control either a mechanical
% or virtual body.
% Can't observe how learning in the model may effect it's evolution in time
% We may miss features about the system
% Brains exist to move muscles

% Solution, build an integrated model. Also give background on OMM system here.

%% End of intro here? Or carry on with a description of the physiology
%% of saccadic eye movements?

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory
feedback, we sought to extend our previous work modelling the oculomotor system.
The existing model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of a
\e{saccadic eye movement};
a fast movement of the eyes which directs the fovea to a region of interest
in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can  be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular system in the body.
%
It is nevertheless behaviourally
interesting, as saccadic eye movements reveal information about decision making at
a subconscious level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context
of building \e{behaving} systems a
necessary part of any model for which the behaviour requires visual attention
and decision making is a realistic mechanism for gathering visual information.
This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within it. It
also follows for \e{intrinsic} reasons. For example, \cite{howard_hand_1997}
showed that visual cues affect reach trajectories and the same group later
demonstrated that reaching affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building the simplest, behaving oculomotor system will
therefore assist future computational neurobehavioural modelling efforts.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a
review, see \cite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral mid-brain \citep{sparks_brainstem_2002} and then through
channels to the motor neurons which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade
target signal in the superior colliculus, it is also possible for
animals to make relatively normal saccades even after the
colliculus has been ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the
cerebellum which are known to participate in the specification, dynamics
and adaptation of saccadic eye movements
\citep{kleine_saccade-related_2003,takagi_effects_1998}.

@@

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade \citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

% Section - description of how the model fits together - which
% components it's made from.
\subsection{Model development framework}

% Say the \ccg~model was implemented in matlab-brahms & this one is SpineML.
The \ccg~model was originally developed to run on the BRAHMS model
execution framework
\citep{mitchinson_brahms:_2010,mitchinson_brahms_2015}. To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
\citep{mitchinson_brahms:_2010}. The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

@@

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted
inputs to the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}
$w_i^{act}$ and $w_i^{sh}$ are, respectively, the weights
of the $i^{th}$ activation or shunting connection; $x_i^{act}$
and $x_i^{sh}$ are the signals input to the
activation and shunting connections.
%
% That's the components described. Next up is to have a figure of the
% activity for representative input to the model.
%

@@

The mapping from the retinotopic co-ordinates in the brain to rotational
co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of
striate cortex [visual stimulus to electrophysiological response---\cite{daniel_representation_1961, talbot_physiological_1941}]
and by \cite{ottes_visuomotor_1986} for superior colliculus data
[electrophysiological SC stimulus to eye movement
response---\cite{robinson_eye_1972}].
% Ottes et al. fit to stimulus-to-movement data, and give a
% movement-to-collicular-coordinate mapping in the paper.
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar coordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar coordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian co-ordinates of the neural map. We use $\theta_{x}^{t}$ and $\theta_{y}^{t}$
to give Euler rotations for the retinal target stimulus. Note also that
the form of Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

@@

% Section - description of new brainstem model
\subsection{Brainstem model}

@@

% Section - description of biomechanical model
\subsection{biomechanical eye}

% Possibly introductory paragraph here.
The output signals of the brainstem are used to drive the biomechanical model.
The latter is not only used to get tangible feedback on the simulated saccades
including motion trajectories, but adds one more modelling dimension related to
the inertial properties of the eye plant including muscle properties.

% EYE MODEL DESCRIPTION HERE
The biomechanical eye model, implemented using the OpenSim framework
\citep{seth_opensim:_2011}, is anatomically represented by a sphere of
uniform mass distribution. The diameter of the eye is 24 mm for adults,
with small variations between individuals; the mass of the eye is 7.5 grams.
The eyeball is actuated by six extraocular muscles
(EOMs). The EOMs are arranged in three pairs forming a cone inside the
orbit with the apex being located inside the cranium in a tendonous
ring called the annulus of Zinn. An important feature of the
oculomotor system which greatly affects its overall behavior is the
existence of dynamic EOM pulleys. Their role is to guide the pivot
point of the EOMs. In our model, a pulley for each EOM has been
modeled by a point on the orbit whose location depends on the current
eye orientation.

%EOM dynamics
The force applied by EOMs is controlled by an excitatory signal
supplied by motoneurons in the brainstem. The neural drive to produce a saccadic
eye movement can be characterized by a pulse component to overcome the
viscoelasticity of the orbital plant, a step component to stabilize the eye in
the new position, and a slide component that models the gradual transition
between the pulse and step.

@@

%Orbital fat force
Passive forces due to the fatty tissues inside the eye orbit also
affect eye dynamics. Their role is critical in eliminating the
influence of head and body movements. We incorporated a custom torque, $\mathbf{t}$,
which acts like a rotational spring-damper apparatus, resisting
eyeball movements. It has elastic and viscous properties governed by
$\mathbf{t} = -K\mathbf{R}-C\mathbf{U}$ where $\mathbf{R}$ is the
eye's orientation and $\mathbf{U}$ is its angular velocity. $K$ and
$C$ are constants. A fuller description of the biomechanical model
can be found in \cite{papapavlou_physics-based_2014}.

% Section - description of the integration. scientific and
% technical. Already
\subsection{Integrating the models and closing the loop}

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a suitable
time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model
there is no such mechanism to reduce activity in SC\_deep, and elsewhere.
Although a successful, accurate saccade towards a target luminance will
remove the excitation
which caused the activity in SC\_deep by bringing the target luminance
within the masked,
foveal region, the activity in SC decays too slowly to avoid
additional saccadic movements. We found it necessary to hypothesize
an inhibitory feedback mechanism from the SBG to the brain model.
This is shown in Fig.~\ref{sbg}, which indicates how the output from
the inhibitory burst neurons (IBN) of the SBG model are used to feed
back an inhibitory signal to the SC\_deep, thalamus and FEF populations
in the brain model, resetting them ready for the next saccade.
% Might have to justify the above a bit more.

>>>>

@@

The field of computational neuroscience has provided
many \e{systems models} of the brain \citep{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}.
We refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain. Whilst they differ in scale and complexity, these models
all seek to
describe the fundamental mechanisms behind common animal behaviours
such as locomotion, threat evasion, reaching or feeding. However,
none of the models cited here actually reproduce these
behaviours. In each case, the activity in a certain population of
neurons is taken to be representative of a behavioural outcome.
In some cases, it \e{is} reasonable to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a basal
ganglia model
\citep{nambu_discharge_1990,kuhn_event-related_2004}. The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. To validate the model, the
error rates which it generates
could be compared with experimentally determined error rates in
primate subjects. We refer
to this as an \e{output assumption model} because the output is assumed
to signify behaviour. (An \e{input assumption model} assumes that sensory
input produces some particular form of neural activity in an input
population of the model.)

However,
we may be interested in reproducing accurate simulated \e{trajectories},
in order to find out how degradation of parts of the model affect
movement. In Parkinson's Disease, degradation of the
dopamine neurons originating in the substantia nigra pars compacta (SNc)
causes diskinesia \citep{galvan_pathophysiology_2008}, as well as
abnormal network activity in
the basal ganglia \citep{brown_dopamine_2001,mccarthy_striatal_2011}.
Sufferers of the disease would be
expected to produce abnormal decision-making \e{and} movement trajectories
in a reach-to-the-correct-target task such as the one described in
\cite{james_target-distractor_2017}.
A model which
sought to explore in detail the effects of the SNc degradation both on
the decision making \e{and} on the movement dynamics would need a
physically accurate virtual arm, as well as physically realistic sensory
input for the brain. This is no less than a complete model of those sections
of the brain and body which act to fulfil the task. Such a modelling
effort, if successful, would result in a virtual robot capable of expressing
behaviour \e{in response to sensory input from its environment}. This would
represent a paradigm shift in the field of computational neuroscience
worthy of the new name of \e{computational neurobehaviour}.

In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory
feedback, we sought to extend our previous work modelling the oculomotor system.
The existing model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of a
\e{saccadic eye movement};
a fast movement of the eyes which directs the fovea to a region of interest
in the field of view.
The oculomotor system is an excellent candidate for modelling because
its movements can  be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular system in the body.
It is nevertheless behaviourally
interesting, as saccadic eye movements reveal information about decision making at
a subconscious level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context
of building \e{behaving} systems a
necessary part of any model for which the behaviour requires visual attention
and decision making is a realistic mechanism for gathering visual information.
This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within it. It
also follows for \e{intrinsic} reasons. For example, \cite{howard_hand_1997}
showed that visual cues affect reach trajectories and the same group later
demonstrated that reaching affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building the simplest, behaving oculomotor system will
therefore assist future computational neurobehavioural modelling efforts.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a
review, see \cite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral mid-brain \citep{sparks_brainstem_2002} and then through
channels to the motor neurons which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
Although both cortical and subcortical paths produce a saccade
target signal in the superior colliculus, it is also possible for
animals to make relatively normal saccades even after the
colliculus has been ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the
cerebellum which are known to participate in the specification, dynamics
and adaptation of saccadic eye movements
\citep{kleine_saccade-related_2003,takagi_effects_1998}.

@@

The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade \citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

\subsection{Model development framework}

The \ccg~model was originally developed to run on the BRAHMS model
execution framework
\citep{mitchinson_brahms:_2010,mitchinson_brahms_2015}. To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
\citep{mitchinson_brahms:_2010}. The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

@@

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted
inputs to the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}
$w_i^{act}$ and $w_i^{sh}$ are, respectively, the weights
of the $i^{th}$ activation or shunting connection; $x_i^{act}$
and $x_i^{sh}$ are the signals input to the
activation and shunting connections.

@@

The mapping from the retinotopic co-ordinates in the brain to rotational
co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of
striate cortex [visual stimulus to electrophysiological response---\cite{daniel_representation_1961, talbot_physiological_1941}]
and by \cite{ottes_visuomotor_1986} for superior colliculus data
[electrophysiological SC stimulus to eye movement
response---\cite{robinson_eye_1972}].
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar coordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar coordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian co-ordinates of the neural map. We use $\theta_{x}^{t}$ and $\theta_{y}^{t}$
to give Euler rotations for the retinal target stimulus. Note also that
the form of Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

@@

\subsection{Brainstem model}

@@

\subsection{biomechanical eye}

The output signals of the brainstem are used to drive the biomechanical model.
The latter is not only used to get tangible feedback on the simulated saccades
including motion trajectories, but adds one more modelling dimension related to
the inertial properties of the eye plant including muscle properties.

The biomechanical eye model, implemented using the OpenSim framework
\citep{seth_opensim:_2011}, is anatomically represented by a sphere of
uniform mass distribution. The diameter of the eye is 24 mm for adults,
with small variations between individuals; the mass of the eye is 7.5 grams.
The eyeball is actuated by six extraocular muscles
(EOMs). The EOMs are arranged in three pairs forming a cone inside the
orbit with the apex being located inside the cranium in a tendonous
ring called the annulus of Zinn. An important feature of the
oculomotor system which greatly affects its overall behavior is the
existence of dynamic EOM pulleys. Their role is to guide the pivot
point of the EOMs. In our model, a pulley for each EOM has been
modeled by a point on the orbit whose location depends on the current
eye orientation.

The force applied by EOMs is controlled by an excitatory signal
supplied by motoneurons in the brainstem. The neural drive to produce a saccadic
eye movement can be characterized by a pulse component to overcome the
viscoelasticity of the orbital plant, a step component to stabilize the eye in
the new position, and a slide component that models the gradual transition
between the pulse and step.

@@

Passive forces due to the fatty tissues inside the eye orbit also
affect eye dynamics. Their role is critical in eliminating the
influence of head and body movements. We incorporated a custom torque, $\mathbf{t}$,
which acts like a rotational spring-damper apparatus, resisting
eyeball movements. It has elastic and viscous properties governed by
$\mathbf{t} = -K\mathbf{R}-C\mathbf{U}$ where $\mathbf{R}$ is the
eye's orientation and $\mathbf{U}$ is its angular velocity. $K$ and
$C$ are constants. A fuller description of the biomechanical model
can be found in \cite{papapavlou_physics-based_2014}.

\subsection{Integrating the models and closing the loop}

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a suitable
time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model
there is no such mechanism to reduce activity in SC\_deep, and elsewhere.
Although a successful, accurate saccade towards a target luminance will
remove the excitation
which caused the activity in SC\_deep by bringing the target luminance
within the masked,
foveal region, the activity in SC decays too slowly to avoid
additional saccadic movements. We found it necessary to hypothesize
an inhibitory feedback mechanism from the SBG to the brain model.
This is shown in Fig.~\ref{sbg}, which indicates how the output from
the inhibitory burst neurons (IBN) of the SBG model are used to feed
back an inhibitory signal to the SC\_deep, thalamus and FEF populations
in the brain model, resetting them ready for the next saccade.

The output signals from the six channels of the SBG were connected to
the six motoneuron inputs of the biomechanical eye. The signal was
normalised; a value of 1 meaning that all the motoneurons in the
output population were firing at their maximum rate and the force
exerted by the relevant extraocular muscle was maximal. Channels
innervated extraocular muscles as follows: Up: superior rectus; Down:
inferior rectus; Right: medial rectus; Left: lateral rectus; Z+:
superior oblique; Z-: inferior oblique. Because the medial rectus
induces a rightward rotation of the eye, our single virtual eye is
a \e{left} eye. The OpenSim implementation of the biomechanical eye
was `wrapped' (in the software sense) in a BRAHMS component. This made
it possible to integrate the OpenSim model into the BRAHMS
framework. The wrapper ensured that the input and output signals were
correctly transferred and, importantly, handled the disparity in the
solver timesteps used in the OpenSim model (25~ms) and the neural
model (1~ms). This was achieved by having the BRAHMS wrapper create a
separate thread to run the OpenSim model. The BRAHMS wrapper component
was called on each 1 ms timestep, receiving the instantaneous
activations from the motoneurons in the SBG. These activations, and
the current simulation time, were written into a shared memory area,
accessible by the OpenSim thread. Running independently, the OpenSim
thread would update its inputs (using the most recent values in the
shared memory area) whenever the simulation time had increased by 25
ms. It would then recompute its outputs (the rotational state of the
eye) and write these into the same shared memory. The BRAHMS wrapper
would update its outputs whenever they were changed in the shared
memory by the OpenSim thread. A direct connection of the six outputs
of the BRAHMS eye model component to the six inputs of the
worldDataMaker BRAHMS component was specified in
the \stob~external.xsl file.

The eye model outputs its rotational state at each
timestep. The rotational state is used to compute the view of the
world in the eye's frame of reference. To simplify
the calculation, the luminances exist on a spherical surface at the
centre of which is the eye. A hand-coded BRAHMS component called
worldDataMaker computes the projection of the luminances into the
eye's frame of reference and then converts this representation into a
retinotopic map to pass into the brain model. The input to the brain
model is thus able to change continuously, on every timestep, rather
than in a step-wise fashion when a saccade occurs, as in the \ccg~model.
